---
title: "cfbResumeR_setup"
author: "Bo Han"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rvest)
library(dplyr)
library(tidyverse)
```

#### Win Loss Records Against Ranked Opponents
```{r}
url = "https://betiq.teamrankings.com/college-football/betting-trends/win-loss-records-vs-ranked-opponent/?season=2024"
page = read_html(url)

team = page %>%
  html_nodes("td.nowrap") %>% html_text(trim=T)
rk_wl = page %>%
  html_nodes(".nowrap+ td") %>% html_text(trim=T)

ranked2024 = data.frame(Team=team, Ranked_WinLoss=rk_wl)
```

```{r}
url = "https://betiq.teamrankings.com/college-football/betting-trends/win-loss-records-vs-ranked-opponent/"
page = read_html(url)

team = page %>%
  html_nodes("td.nowrap") %>% html_text(trim=T)
rk_wl = page %>%
  html_nodes(".nowrap+ td") %>% html_text(trim=T)

ranked2025 = data.frame(Team=team, Ranked_WinLoss=rk_wl)
```

#### Win Loss Records Against Conference Opponents
```{r}
url = "https://betiq.teamrankings.com/college-football/betting-trends/win-loss-records-conference-games/?season=2024"
page = read_html(url)

team = page %>%
  html_nodes("td.nowrap") %>% html_text(trim=T)
conf_wl = page %>%
  html_nodes(".nowrap+ td") %>% html_text(trim=T)

conference2024 = data.frame(Team=team, Conference_WinLoss=conf_wl)
```

```{r}
url = "https://betiq.teamrankings.com/college-football/betting-trends/win-loss-records-conference-games/"
page = read_html(url)

team = page %>%
  html_nodes("td.nowrap") %>% html_text(trim=T)
conf_wl = page %>%
  html_nodes(".nowrap+ td") %>% html_text(trim=T)

conference2025 = data.frame(Team=team, Conference_WinLoss=conf_wl)
```

#### Win Loss Records for Regular Season
```{r}
url = "https://betiq.teamrankings.com/college-football/betting-trends/win-loss-records-regular-season/?season=2024"
page = read_html(url)

team = page %>%
  html_nodes("td.nowrap") %>% html_text(trim=T)
reg_wl = page %>%
  html_nodes(".nowrap+ td") %>% html_text(trim=T)

season2024 = data.frame(Team=team, RegSeason_WinLoss=reg_wl)
```

```{r}
url = "https://betiq.teamrankings.com/college-football/betting-trends/win-loss-records-regular-season/"
page = read_html(url)

team = page %>%
  html_nodes("td.nowrap") %>% html_text(trim=T)
reg_wl = page %>%
  html_nodes(".nowrap+ td") %>% html_text(trim=T)

season2025 = data.frame(Team=team, RegSeason_WinLoss=reg_wl)
```

#### Join Tables
```{r}
winloss_2024 = ranked2024 %>%
  left_join(conference2024, by="Team") %>%
  left_join(season2024, by="Team")

winloss_2025 = ranked2025 %>%
  left_join(conference2025, by="Team") %>%
  left_join(season2025, by="Team")
```

#### Standardize Team Names
```{r}
name_map = c(
  "Ohio St" = "Ohio State",
  "Penn St" = "Penn State",
  "Iowa St" = "Iowa State",
  "Kansas St" = "Kansas State",
  "Michigan St" = "Michigan State",
  "Mississippi St" = "Mississippi State",
  "Oklahoma St" = "Oklahoma State",
  "Oregon St" = "Oregon State",
  "Utah St" = "Utah State",
  "Kent St" = "Kent State",
  "Boise St" = "Boise State",
  "Fresno St" = "Fresno State",
  "San Diego St" = "San Diego State",
  "San Jose St" = "San José State",
  "Florida St" = "Florida State",
  "N Illinois" = "Northern Illinois",
  "E Carolina" = "East Carolina",
  "C Michigan" = "Central Michigan",
  "W Michigan" = "Western Michigan",
  "W Kentucky" = "Western Kentucky",
  "S Florida" = "South Florida",
  "N Texas" = "North Texas",
  "Middle Tenn" = "Middle Tennessee",
  "Mississippi" = "Ole Miss",
  "Miami OH" = "Miami (OH)",
  "Coastal Car" = "Coastal Carolina",
  "Florida Intl" = "Florida International",
  "J Madison" = "James Madison",
  "UMass" = "Massachusetts",
  "Ball St" = "Ball State",
  "Georgia So" = "Georgia Southern",
  "Jacksonville St" = "Jacksonville State",
  "Arkansas St" = "Arkansas State",
  "Arizona St" = "Arizona State",
  "Georgia St" = "Georgia State",
  "New Mexico St" = "New Mexico State",
  "Washington St" = "Washington State",
  "Kennesaw St" = "Kennesaw State",
  "Colorado St" = "Colorado State",
  "S Alabama" = "South Alabama"
)

winloss_2024 = winloss_2024 %>%
  mutate(Team=recode(Team, !!!name_map))

winloss_2025 = winloss_2025 %>%
  mutate(Team = recode(Team, !!!name_map))
```

#### Separate Wins and Losses
```{r}
split_record = function(df, col){
  df %>%
    separate({{col}}, into=c("Wins", "Losses"), sep="-", convert=T)
}
```

```{r, warning=F}
winloss_2024 = winloss_2024 %>%
  separate(Ranked_WinLoss, into=c("Ranked_Wins", "Ranked_Losses"), sep="-",
           convert=T, remove=F) %>%
  separate(Conference_WinLoss, into=c("Conf_Wins", "Conf_Losses"), sep="-",
           convert=T, remove=F) %>%
  separate(RegSeason_WinLoss, into=c("RegSeason_Wins", "RegSeason_Losses"), sep="-",
           convert=T, remove=F)

winloss_2025 = winloss_2025 %>%
  separate(Ranked_WinLoss, into=c("Ranked_Wins", "Ranked_Losses"), sep="-",
           convert=T, remove=F) %>%
  separate(Conference_WinLoss, into=c("Conf_Wins", "Conf_Losses"), sep="-",
           convert=T, remove=F) %>%
  separate(RegSeason_WinLoss, into=c("RegSeason_Wins", "RegSeason_Losses"), sep="-",
           convert=T, remove=F)
```

#### Convert Wins/Losses into Win Percentages
```{r}
winloss_2024 = winloss_2024 %>%
  mutate(Ranked_WinPct = Ranked_Wins/(Ranked_Wins+Ranked_Losses),
         Conf_WinPct = ifelse(is.na(Conf_Wins),
                              0,
                              Conf_Wins/(Conf_Wins+Conf_Losses)),
         RegSeason_WinPct = RegSeason_Wins/(RegSeason_Wins+RegSeason_Losses)
  ) %>%
  select("Team", "Ranked_WinPct", "Conf_WinPct", "RegSeason_WinPct",
         "RegSeason_WinLoss")

winloss_2025 = winloss_2025 %>%
  mutate(Ranked_WinPct = Ranked_Wins/(Ranked_Wins+Ranked_Losses),
         Conf_WinPct = ifelse(is.na(Conf_Wins),
                              0,
                              Conf_Wins/(Conf_Wins+Conf_Losses)),
         RegSeason_WinPct = RegSeason_Wins/(RegSeason_Wins+RegSeason_Losses)
  ) %>%
  select("Team", "Ranked_WinPct", "Conf_WinPct", "RegSeason_WinPct",
         "RegSeason_WinLoss")
```

#### Season Stats For Teams
```{r}
teams2024 = read.csv("2024_season.csv")
teams2024 = teams2024 %>%
  pivot_wider(names_from=StatName,
              values_from=StatValue
  )

teams2025 = read.csv("2025_season.csv")
teams2025 = teams2025 %>%
  pivot_wider(names_from=StatName,
              values_from=StatValue
  )
```

#### Create Model Features
```{r}
teams2024 = teams2024 %>%
  mutate(yards_diff_pg = (totalYards-totalYardsOpponent)/games,
         pass_yards_diff_pg = (netPassingYards-netPassingYardsOpponent)/games,
         rush_yards_diff_pg = (rushingYards-rushingYardsOpponent)/games,
         td_diff_pg =
           (passingTDs+rushingTDs-passingTDsOpponent-rushingTDsOpponent)/games,
         turnover_diff_pg = (turnoversOpponent-turnovers)/games,
         interception_diff_pg = (interceptions-interceptionsOpponent)/games,
         fumble_diff_pg = (fumblesRecovered-fumblesLost)/games,
         sack_diff_pg = (sacks-sacksOpponent)/games,
         tfl_diff_pg = (tacklesForLoss-tacklesForLossOpponent)/games,
         penalties_pg = penalties/games,
         penalty_yards_pg = penaltyYards/games,
         possession_diff = possessionTime-possessionTimeOpponent
  )

teams2025 = teams2025 %>%
  mutate(yards_diff_pg = (totalYards-totalYardsOpponent)/games,
         pass_yards_diff_pg = (netPassingYards-netPassingYardsOpponent)/games,
         rush_yards_diff_pg = (rushingYards-rushingYardsOpponent)/games,
         td_diff_pg =
           (passingTDs+rushingTDs-passingTDsOpponent-rushingTDsOpponent)/games,
         turnover_diff_pg = (turnoversOpponent-turnovers)/games,
         interception_diff_pg = (interceptions-interceptionsOpponent)/games,
         fumble_diff_pg = (fumblesRecovered-fumblesLost)/games,
         sack_diff_pg = (sacks-sacksOpponent)/games,
         tfl_diff_pg = (tacklesForLoss-tacklesForLossOpponent)/games,
         penalties_pg = penalties/games,
         penalty_yards_pg = penaltyYards/games,
         possession_diff = possessionTime-possessionTimeOpponent
  )
```

#### Add Win Percentages to Team Stats Dataset
```{r}
teams2024 = teams2024 %>%
  left_join(winloss_2024, by="Team") %>%
  drop_na()

teams2025 = teams2025 %>%
  left_join(winloss_2025, by="Team") %>%
  drop_na()
```

#### Standardize All Model Features
```{r}
teams2024 = teams2024 %>%
  mutate(across(c(yards_diff_pg,
                  pass_yards_diff_pg,
                  rush_yards_diff_pg,
                  td_diff_pg,
                  turnover_diff_pg,
                  interception_diff_pg,
                  fumble_diff_pg,
                  sack_diff_pg,
                  tfl_diff_pg,
                  penalties_pg,
                  penalty_yards_pg,
                  possession_diff,
                  Ranked_WinPct,
                  Conf_WinPct,
                  RegSeason_WinPct
                ),
                ~ scale(.)[,1]
  ))

teams2025 = teams2025 %>%
  mutate(across(c(yards_diff_pg,
                  pass_yards_diff_pg,
                  rush_yards_diff_pg,
                  td_diff_pg,
                  turnover_diff_pg,
                  interception_diff_pg,
                  fumble_diff_pg,
                  sack_diff_pg,
                  tfl_diff_pg,
                  penalties_pg,
                  penalty_yards_pg,
                  possession_diff,
                  Ranked_WinPct,
                  Conf_WinPct,
                  RegSeason_WinPct
                ),
                ~ scale(.)[,1]
  ))
```

#### Weigh Features to Create Resume Score
```{r}
teams2024 = teams2024 %>%
  mutate(score =
           0.10*(pass_yards_diff_pg+rush_yards_diff_pg+td_diff_pg) +
           0.10*(interception_diff_pg+fumble_diff_pg+sack_diff_pg+tfl_diff_pg) +
           0.05*(turnover_diff_pg+possession_diff) -
           0.05*(penalties_pg+penalty_yards_pg) +
           0.50*(Ranked_WinPct) +
           0.15*(RegSeason_WinPct) +
           0.05*(Conf_WinPct)
  )

teams2025 = teams2025 %>%
  mutate(score =
           0.10*(pass_yards_diff_pg+rush_yards_diff_pg+td_diff_pg) +
           0.10*(interception_diff_pg+fumble_diff_pg+sack_diff_pg+tfl_diff_pg) +
           0.05*(turnover_diff_pg+possession_diff) -
           0.05*(penalties_pg+penalty_yards_pg) +
           0.50*(Ranked_WinPct) +
           0.15*(RegSeason_WinPct) +
           0.05*(Conf_WinPct)
  )
```

#### Find Playoff Picture
```{r}
teams2024 %>%
  arrange(desc(score)) %>%
  mutate(Rank=row_number(),
         Record=RegSeason_WinLoss) %>%
  select(Team, Conference, Rank, Record) %>%
  head(12)
```

```{r}
teams2025 %>%
  arrange(desc(score)) %>%
  mutate(Rank=row_number(),
         Record=RegSeason_WinLoss) %>%
  select(Team, Conference, Rank, Record) %>%
  head(12)
```

#### Actual 2024 and 2025 Playoff Teams
```{r}
playoff24_tms = c("Oregon", "Georgia", "Boise State"," Arizona State", "Texas",
                "Penn State", "Notre Dame", "Ohio State", "Tennessee", "Indiana",
                "SMU", "Clemson")

playoff25_tms = c("Indiana", "Ohio State", "Georgia", "Texas Tech", "Oregon",
                  "Ole Miss", "Texas A&M", "Oklahoma", "Alabama", "Miami", "Tulane",
                  "James Madison")
```

